HO unification from object language to meta language

### 2024-03-30 paper/main.tex
add notations
alternative list of maybe-eta examples
tex wip

\noindent
When detailing examples we write links as equations between two
terms under a context.
The equality sign is subscripted with
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-beta (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

@@

A second example, showing the activation of a link when the \rhs is no more a
\maybeeta, is given in \cref{sec:invariant1}, since we need to work with
variables used with different arities.

TODO: example for case 2: $\lambda x. \lambda y. F \appsep y \appsep x = G, F = \lambda x. \lambda y. a$

A second way to progress \linketa, that we call \linketa deduplication, is when
\linkStore contains two \linketa $l_1$ and $l_2$ with a \lhs having the same
variable address. Let the \lhs of $l_1$ be \elpiIn{uva U R} and the \lhs of
$l_2$ be \elpiIn{uva V S}, then, by \cref{inv:uvaarity}, \elpiIn{R}
and \elpiIn{S} have same scope. Let $t$ be the term obtained by replacing all
each name \elpiIn{S}$_i$ in the \rhs of $l_1$ with \elpiIn{R}$_i$, $t$ is
unified with the \rhs of $l_2$ and one of the two links between $l_1$ and $l_2$
is removed from \linkStore.

TODO: example for this: $\lambda x. \lambda y. F \appsep y \appsep x = X,
\lambda x. \lambda y. F \appsep y \appsep x = Y$

@@

% In this subsection we show two complete runs of the extended version of \Ue.

% \paragraph{Example 1}
% \begin{center}
%   \elpiIn{flam x\ flam y\ fapp[fuva A, y, x] = flam x\ flam y\ x}
% \end{center}
% % $$\flambda xy.F \appsep y \appsep x = \lambda xy.x$$
% \begin{itemize}
%   \item After compilation: terms are $X0 = \lambda c0.\lambda c1.c0$ with links
%         $\linketaM{c0}{X1_{c0}}{\lambda c1.X2_{c1~c0}}, \linketaM{}{X0}{\lambda
%         c0.X1_{c0}}$ and mappings: $\mapping{C0}{X2}{2}$
%   \item After unification $X0$ is set to $\lambda c0.\lambda c1.c0$
%   \item This unification wakes up the second link, since its \lhs is now
%         instantiated. $X0$ is unified with its \rhs, which assigns $X1_{c0} :=
%         \lambda c1.c0$
%   \item The instantiation of $X1$ wakes the first link, and $\lambda c1.c0$ is
%         unified with $\lambda c1.X2_{c1~c0}$, setting $X2 _{c1~c2} := c2$ by
%         pruning.
%   \item All the links have been solved and the decompilation, assigns the \Fo{}
%         variable \elpiIn{A} to $\lambda xy.y$ 
% \end{itemize}

% \paragraph{Example 2}
% \begin{center}
%   \elpiIn{flam x\ fapp[f, fapp[fuva F, x], x] = fapp[f, flam x\ x]}
% \end{center}
% % $$\flambda xy.F \appsep y \appsep x = \lambda xy.x$$
% \begin{itemize}
%   \item After compilation: terms are $X0 = f~\lambda c0.c0$ with links
%         $\linketaM{}{X0}{\lambda c0.(f~X1_{c0}~c0)}$ and mappings:
%         $\mapping{C0}{X1}{1}$
%   \item After unification $X0$ is set to $f~\lambda c0.c0$
%   \item This unification wakes up the link, since its \lhs is now
%         instantiated. $X0$ is unified with its \rhs, which assigns $X1_{c0} :=
%         \lambda c1.c0$
%   \item The instantiation of $X1$ wakes the first link, and $\lambda c1.c0$ is
%         unified with $\lambda c1.X2_{c1~c0}$, setting $X2 _{c1~c2} := c2$ by
%         pruning.
%   \item All the links have been solved and the decompilation, assigns the \Fo{}
%         variable \elpiIn{A} to $\lambda xy.y$ 
% \end{itemize}

>>>>

\noindent
When variables $x$ and $y$ can occur in term $t$ we shall write
$t_{xy}$ to stress this fact.

We write $\sigma = \{ A_{xy} \mapsto y\}$ for the assignment
\elpiIn{abs x\abs y\y} and $\sigma = \{ A \mapsto \lambda x.\lambda y.y\}$
for \elpiIn{lam x\lam y\y}.

When detailing examples we write links as equations between two
terms under a context.
The equality sign is subscripted with
kind of \elpiIn{baselink}. For example $\linkbetaM{x}{A_x}{F_x~a}$ corresponds to:
\begin{elpicode}
abs x\ val (link-beta (uva A [x]) (app[uva F [x],con "a"]))
\end{elpicode}

@@

\begin{center}
  \begin{tabular}{lll}
    Term & Status & Evidence \\\hline
    $\lambda x. f \appsep A_x$ & $\in\maybeeta$ & $\sigma = \{~ A_x \mapsto x ~\}$ \\
    $\lambda x. f \appsep x \appsep A_x$ & $\not\in\maybeeta$ &\\
    $\lambda x. \lambda y. f \appsep A_x \appsep B_{y x}$ & $\in\maybeeta$ & $\sigma = \{~ A_x \mapsto x,~ B_{yx} \mapsto y ~\}$\\
    $\lambda x. f \appsep A_x \appsep x$ & $\in\maybeeta$ & $\sigma = \{~ A_x \mapsto a ~\}$
  \end{tabular}
\end{center}

@@

A second example, showing the activation of a link when the \rhs is no more a
\maybeeta, is given in \cref{sec:invariant1}, since we need to work with
variables used with different arities. This example represent the run of the
unification problems proposed at \cref{sec:basic-comp-limitations}

Another, way to progress \linketa, that we call \linketa deduplication, is when
\linkStore contains two \linketa $l_1$ and $l_2$ with a \lhs having the same
variable address. Let the \lhs of $l_1$ be \elpiIn{uva U R} and the \lhs of
$l_2$ be \elpiIn{uva V S}, then, by \cref{inv:uvaarity}, \elpiIn{R}
and \elpiIn{S} have same scope. Let $t$ be the term obtained by replacing all
each name \elpiIn{S}$_i$ in the \rhs of $l_1$ with \elpiIn{R}$_i$, $t$ is
unified with the \rhs of $l_2$ and one of the two links between $l_1$ and $l_2$
is removed from \linkStore.

A very basic example of \linketa deduplication, is given below:

\printAll
  {{\elpiIn{flam x\ fapp [fuva X, x] == flam x\ fapp [fuva X, x]}}}
  {{E0 == E2}}
  {{\mapping{X}{X'}{1}}}
  {{\linketaM{}{E2}{\lambda x.X'_{x}},
    \linketaM{}{E0}{\lambda x.X'_{x}}}}

TODO: explain

@@
